Public Support for Democracy, Regime Change, and the Importance of Incorporating Measurement Uncertainty

### 2020-12-29 paper/dcpo_demsupport.Rmd
work toward revising introduction

One classic but disputable question in comparative politics is the relationship between mass democratic support and countries' level of democracy.
A long-standing and widely accepted theory is that the level of mass democratic support determines the stability of a democratic system [e.g., @lipset_social_1959; @easton1965systems; @norris_democratic_2011], but empirical findings remain inconclusive [@inglehart_modernization_2005; @fails_changing_2010; @qi_how_2011; for a review of this literature, see @Claassen2020a].

One reason for the mixed results is the difficulties in measuring comparative public democratic support.
Public support for democracy is a latent variable, and the incorrect measurement of it may limit inferences about relationship between public opinion and democratic institutional development.
However, the sparsity and incomparability of public opinion data across countries over time have always been perplexing comparative studies for a long time [@Solt2020a].
Recently, a few pioneering studies have contributed to comparative public opinion studies using advanced latent variable measurement [@Caughey2019; @Claassen2019; @Solt2020a].

A prominent study by @Claassen2020a applied an advanced Bayesian Item Response Model to measure dynamic public democratic support for 135 countries for up to 29 years.
It found that mass support had a positive impact on democratic change, especially the endurance of democracy.
The analysis models he used do not account, though, for measurement uncertainty.
The latent constructs are inherently accompanied by a quantity  of measurement uncertainty due to the unobservability of variables.
Left unacknowledged, with a strong assumption that latent variables are measured perfectly, measurement error could distort both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].

>>>>

One classic but unresolved question in comparative politics is the relationships between mass democratic support and countries' levels of democracy.
A long-standing and widely accepted theory is that the extent of mass democratic support determines the stability of a democratic system [e.g., @lipset_social_1959; @easton1965systems; @norris_democratic_2011], but empirical findings remain inconclusive [@inglehart_modernization_2005; @fails_changing_2010; @qi_how_2011; for a review of this literature, see @Claassen2020a].

One important reason for these mixed results is the difficulty in measuring comparative public democratic support.
Public support for democracy cannot be directly observed, and its incorrect measurement will limit inferences about the relationship between public opinion and democratic institutional development.
However, the survey data available across countries and over time on support for democracy--or indeed nearly any topic in public opinion---are sparse and incomparable, greatly hindering the broadly comparative research needed to reach sound conclusions.
Recently, a few pioneering studies have sought to overcome the hurdle of sparse and incomparable data and further broadly comparative research by developing latent variable measurement models of public opinion [see @Caughey2019; @Claassen2019; @Solt2020a].
One prominent study, @Claassen2020a took advantage of this approach to measure dynamic public democratic support for 135 countries for up to 29 years and to then assess its relationship to democratic change.

It found that mass support had a positive impact on democratic change, especially the endurance of democracy.

The analysis models he used do not account, though, for measurement uncertainty.
The latent constructs are inherently accompanied by a quantity  of measurement uncertainty due to the unobservability of variables.
Left unacknowledged, with a strong assumption that latent variables are measured perfectly, measurement error could distort both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].
